---
title: "WGCNA"
output: html_document
date: "2024-06-14"
---

```{r}
library(WGCNA)
library(DESeq2)
library(GEOquery)
library(tidyverse)
library(gridExtra)
```

```{r}
setwd('~/Downloads/')
```


```{r}
# Import data
data <- readr::read_delim("GSE152418_p20047_Study1_RawCounts.txt", delim = "\t")
```
```{r}
# Import metadata
geo_id <- "GSE152418"
gse <- getGEO(geo_id, GSEMatrix = T )
```
```{r}
# Extract phenodata
phenoData <- pData(phenoData(gse[[1]]))
```

```{r}
#subset interesting columns in phenoData
phenoData <- phenoData[,c(1,2,46:50)]
```

```{r}
data <- data %>% 
  gather(key='samples' , value='counts', -ENSEMBLID) %>%
  mutate(samples = gsub('\\.','-', samples)) %>%
  inner_join(.,phenoData, by = c('samples' = 'title')) %>%
  select(1,3,4) %>%
  spread(key= 'geo_accession' , value= 'counts') %>%
  column_to_rownames(var = 'ENSEMBLID')
```

# QC: Outlier Detection
```{r}
gsg <- goodSamplesGenes(t(data))
summary(gsg)
gsg$allOK
```
```{r}
table(gsg$goodGenes)
table(gsg$goodSamples)
```
```{r}
# remove outlier genes
data <-data[gsg$goodGenes == T,]
```

```{r}
# remove outlier samples using PCA
pca <- prcomp(t(data))
```

```{r}
pca.dat <- pca$x
```

```{r}
pca.var <- pca$sdev^2
var.percent <- round(pca.var/sum(pca.var)*100, digits=2)
```

```{r}
pca.dat <- as.data.frame(pca.dat)
```

```{r}
ggplot(pca.dat, aes(PC1, PC2)) +
  geom_point() +
  geom_text(label = rownames(pca.dat))
``` 
```{r}
exclude <- c('GSM4615000', 'GSM4614993', 'GSM4614995')
data.subset <- data[,!(colnames(data) %in% exclude)]
```

#Normalization

```{r}
colData <- phenoData %>% filter(!row.names(.) %in% exclude)
```

```{r}
names(colData)
```

```{r}
names(colData) <- gsub(":ch1","", names(colData))
names(colData) <- gsub("\\s","_", names(colData))
```

```{r}
names(colData)
```

```{r}
all(rownames(colData) %in% colnames(data.subset))
all(rownames(colData) == colnames(data.subset))
```
```{r}
dds <- DESeqDataSetFromMatrix(countData = data.subset , colData = colData, design = ~1)
```

```{r}
# remove all genes with counts less than 15 in 75% of samples
dds75 <- dds[rowSums(counts(dds) >=15) >=24,]
```

```{r}
# Variance stabilization
dds_norm <- vst(dds75)
```

```{r}
#Normailized counts
norm_counts <- assay(dds_norm) %>% t()
```

# Network Construction
## soft thresholding powers are used to transform the similarity matrix into an adjacency matrix in a way that emphasizes stronger correlations while suppressing weaker ones.

```{r}
power <- c(c(1:10), seq(from=12, to= 50, by=2))
```
```{r}
# Network topology analysis
sft <- pickSoftThreshold(norm_counts, powerVector = power, networkType = "signed", verbose = 5)
```
```{r}
sft.data <- sft$fitIndices
```

```{r}
a1 <- ggplot(sft.data, aes(Power, SFT.R.sq, label = Power)) +
  geom_point() +
  geom_text(nudge_y = 0.1) +
  geom_hline(yintercept = 0.8, color = 'red') +
  labs(x = 'Power', y = 'Scale free topology model fit, signed R^2') +
  theme_classic()
```

#The scale-free topology model fit is a key metric used in Weighted Gene Co-expression Network Analysis (WGCNA) to determine the optimal soft-thresholding power for constructing the co-expression network.

```{r}
a1
```
```{r}
# convert matrix to numeric
norm_counts[] <- sapply(norm_counts, as.numeric)
soft_power <- 18
temp_cor <- cor
cor <- WGCNA::cor
```

```{r}
bwnet <- blockwiseModules(norm_counts,
                 maxBlockSize = 14000,
                 TOMType = "signed",
                 power = soft_power,
                 mergeCutHeight = 0.25,
                 numericLabels = FALSE,
                 randomSeed = 1234,
                 verbose = 3)
```
```{r}
cor <- temp_cor
```

# Module Eigengenes

```{r}
module_eigengenes <- bwnet$MEs
head(module_eigengenes)
```

```{r}
table(bwnet$colors)
```
```{r}
plotDendroAndColors(bwnet$dendrograms[[1]], cbind(bwnet$unmergedColors, bwnet$colors),
                    c("unmerged", "merged"),
                    dendroLabels = FALSE,
                    addGuide = TRUE,
                    hang= 0.03,
                    guideHang = 0.05)

```

